The Minimal Retroreflective Microfacet Model

### 2025-12-29 MRM.tex
Missing fullstop in equation.
Plot color improvements

\paragraph*{Jacobian of Back-Vector.}
First we observe that the Jacobian of the change of variable from the back-vector to the light vector $\vecl$ for specular microfacets~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
\begin{equation}
\left\lVert \dfrac{\partial \omega_{\vecb_{r}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{r}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech_{r}(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecbr} \right|} \ .
\end{equation}
We also observe that this is true for refractive microfacets~(Equation (17) of \citet{WalterMicrofacet}):
\begin{equation}
\left\lVert \dfrac{\partial \omega_{\vecb_{t}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{t}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{  \eta_o^2 \left| \vecl, \vech_{t}(\vecvv, \vecl)  \right|}{ \left\lVert \vech_{t}(\vecvv, \vecl) \right\rVert^2}
\end{equation}
This is due to the fact that we have to differentiate with respect to the light direction which is not altered by the back vector.

@@

\paragraph*{Final form.}
It follows from the above back-vector modifications that our MRM retroreflective model takes the trivial form of simply replacing $\vecv$ by $\vecvv$ in the original microfacet BRDF and BTDF expressions:
\begin{equation} \label{eq:retro_microfacet_model}
  %  f_\text{MRM}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
  f_\text{MRM}(\vecv, \vecl) =
  \begin{cases}
    f_{r, \text{MRM}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if \;} (\vdot{\vecv}{\vecn})(\vdot{\vecl}{\vecn}) \geq 0, \\
    f_{t, \text{MRM}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
  \end{cases}
\end{equation}
\begin{figure}[!tb]
  \centering
  \includegraphics[width=0.93\linewidth]{figures/brdf_plots.pdf}
  \caption{Comparison of the retroreflective MRM BRDF (blue) to the regular GGX microfacet BRDF (red) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in \{45^\circ, 80^\circ\}$. The dashed lines are the roughness $0.1$ case, and the solid lines roughness $0.666$. The horizontal green line at $1/\pi$ is the Lambert BRDF.}
  \label{fig:brdf_plots}
\end{figure}

@@

\begin{figure}[!t]
  \begin{center}
    \includegraphics[width=0.99\textwidth]{figures/brdf_theta_L_plots.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{Measured BRDF of retroreflective tape (blue dots) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (red line).}
  \label{fig:data_fit}
\end{figure}

>>>>

\paragraph*{Jacobian of Back-Vector.}
First we observe that the Jacobian of the change of variable from the back-vector to the light vector $\vecl$ for specular microfacets~(Equation~(11) and~(14) of \citet{WalterMicrofacet}) has the same form as the half-vector one:
\begin{equation}
\left\lVert \dfrac{\partial \omega_{\vecb_{r}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{r}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{1}{4 \left| \vdot{\vecl}{\vech_{r}(\vecvv, \vecl)} \right|} = \dfrac{1}{4 \left| \vdot{\vecl}{\vecbr} \right|} \ .
\end{equation}
We also observe that this is true for refractive microfacets~(Equation (17) of \citet{WalterMicrofacet}):
\begin{equation}
\left\lVert \dfrac{\partial \omega_{\vecb_{t}}}{\partial \omega_{\vecl}} \right\rVert = \left\lVert \dfrac{\partial \omega_{\vech_{t}(\vecvv, \vecl)}}{\partial \omega_{\vecl}} \right\rVert = \dfrac{  \eta_o^2 \left| \vecl, \vech_{t}(\vecvv, \vecl)  \right|}{ \left\lVert \vech_{t}(\vecvv, \vecl) \right\rVert^2} \ .
\end{equation}
This is due to the fact that we have to differentiate with respect to the light direction which is not altered by the back vector.

@@

\paragraph*{Final form.}
It follows from the above back-vector modifications that our MRM retroreflective model takes the trivial form of simply replacing $\vecv$ by $\vecvv$ in the original microfacet BRDF and BTDF expressions:
\begin{equation} \label{eq:retro_microfacet_model}
  %  f_\text{MRM}(\vecv, \vecl) = f(\vecvv, \vecl) = \frac{D(\vecb) \, G_{2}(\vecvv,\vecl) \, F(\vecvv, \vecb)}{4 \vdot{\vecvv}{\vecn} \vdot{\vecl}{\vecn}}
  f_\text{MRM}(\vecv, \vecl) =
  \begin{cases}
    f_{r, \text{MRM}}(\vecv, \vecl) = f_r(\vecvv, \vecl)  & \text{if \;} (\vdot{\vecv}{\vecn})(\vdot{\vecl}{\vecn}) \geq 0, \\
    f_{t, \text{MRM}}(\vecv, \vecl) = f_t(\vecvv, \vecl)  & \text{otherwise}.
  \end{cases}
\end{equation}
\begin{figure}[!tb]
  \centering
  \includegraphics[width=0.93\linewidth]{figures/brdf_plots.pdf}
  \caption{Comparison of the retroreflective MRM BRDF (green) to the regular GGX microfacet BRDF (blue) and rough-diffuse EON model (orange), as a function of view direction $\theta_o$ for fixed light directions $\theta_i \in \{45^\circ, 80^\circ\}$. The dashed lines are the roughness $0.1$ case, and the solid lines roughness $0.666$. (The horizontal red line at $1/\pi$ is the Lambert BRDF).}
  \label{fig:brdf_plots}
\end{figure}

@@

\begin{figure}[!t]
  \begin{center}
    \includegraphics[width=0.99\textwidth]{figures/brdf_theta_L_plots.pdf}
  \end{center}
  \vspace{-0.5cm}
  \caption{Measured BRDF of retroreflective tape (black dots) at three different light source angles $\theta_L \in \{15^\circ, 30^\circ, 60^\circ\}$, and the MRM model fit (green dashed line).}
  \label{fig:data_fit}
\end{figure}
